General Physics Formula

### 2016-10-24 GRE Physics Prep/General Physics Formula.tex
A lot of review notes for physics GRE.

@@

This acceleration is due to a net centripetal force on the particle,
with magnitude given by
\begin{align}
    F = \frac{mv^2}{R}
\end{align}
where $m$ is the particle’s mass. The vector quantities $\vec a$ and
$\vec F$ are directed toward the center of curvature of the particle’s
path
\begin{defi}[Normal force]
\nomenclature{Normal force}{\nomrefpage.}
A \textbf{normal force} is the force on a body from a surface against
which the body presses.The normal force is always perpendicular to the
surface.
\end{defi}

@@

The linear acceleration $\vec{a}$ of the point \hl{has both tangential and
radial components}.The tangential component is
\begin{align}
    a_t = \alpha r
\end{align}
where $\alpha$ is the magnitude of the angular acceleration (in radians
per second-squared) of the body.The radial component of $\vec{a}$ is
\begin{align}
    a_r = \frac{v^2}{r} = \omega^2 r
\end{align}
If the point moves in uniform circular motion, the period $T$ of
the motion for the point and the body is
\begin{align}
    T = \frac{2\pi r}{v} = \frac{2\pi}{\omega}
\end{align}
\paragraph{Rotational Kinetic Energy and Rotational Inertia} The
kinetic energy $K$ of a rigid body rotating about a fixed axis is
given by
\begin{align}
    K=\frac{1}{2} I\omega^2
\end{align}
in which $I$ is the rotational inertia of the body, defined as
\begin{align}
    I \equiv \int r^2 \diff m
\end{align}
\paragraph{The Parallel-Axis Theorem} 265
\begin{align}
    I = I_{\text{com}} + M h^2
\end{align}

\paragraph{Work and Rotational Kinetic Energy} 265
\begin{align}
    W = \int_{\theta_i}^{\theta_f} \tau \diff \theta \\
    P = \frac{\diff W}{\diff t} = \tau \omega
\end{align}

@@

\paragraph{Conduction} (\textbf{pp.498 ch18 of \cite{book}})
\textbf{Thermal expansion}
\begin{align}
    & \Delta L = L \alpha \Delta T \\
    & \Delta V = V \beta \Delta T \\
    & \beta = 3 \alpha \\
    & Q = C \Delta T = cm \Delta T
\end{align}
\begin{defi}[Convection]
\nomenclature{Convection}{\nomrefpage.}
    Convection occurs when temperature differences cause an energy
    transfer by motion within a fluid
\end{defi}
\textbf{Radiation}
\begin{align}
    & P_\text{rad} = \sigma \varepsilon AT^4 \\
    & P_\text{abs} = \sigma \varepsilon AT_\text{evn}^4
\end{align}
\textbf{Conduction}
\begin{align}
    P_\text{cond} \equiv \frac{Q}{t} = kA \frac{T_H-T_C}{L} 
\end{align}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{ch18.heat_cond}.png}
\end{figure}

\paragraph{Thermal Process} (\textbf{pp.529 ch19 of \cite{book}})
Thermal process:
\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
        No. & Constant Quaility            & Process Type & Special Results  \\
        \hline
        1   & $p$                          & Isobaric     & $Q = nC_p\Delta T;W=p\Delta V$ \\
        2   & $V$                          & Isothermal   & $Q=W=nRT\ln(V_f/V_i); \Delta E_\text{int}=0$  \\
        3   & $pV^\gamma$, $TV^{\gamma-1}$ & Adiabatic    & $W=-\Delta E_\text{int}$ \\
        4   & $V$                          & Isochoric    & $Q=\Delta E_\text{int}=nC_V \Delta T;W=0$ \\
        \hline
    \end{tabular}
\end{table}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{ch18.therm_pro}.png}
\end{figure}
Isothermal Process:
\begin{align}
    W = Nk T \ln\frac{V_f}{V_i}
\end{align}
\textbf{Temperature and Kinetic Energy}
\begin{align}
    & K_\text{avg} = \frac{3}{2}kT \text{, per molecule} \\
    & \lambda = \frac{1}{\sqrt{2} \pi d^2 N/V} 
\end{align}
\paragraph{Molar Specific Heats}
\begin{align}
    & C_V = \frac{3}{2} R \\
    & C_P = C_V + R \\
    & \Delta E_\text{int} = N C_V \Delta T \\
    & C_V = \frac{f}{2} k \\
    & \gamma \equiv \frac{C_P}{C_V}
\end{align}

@@

\begin{align}
    & \mathcal{E} = \frac{\diff W}{\diff q}
\end{align}

@@

@@

@@

\paragraph{Relativity} (\textbf{pp.1048 ch37 of \cite{book}})
\begin{align}
    \Delta t &= \gamma \Delta t_0 \\
    L &= L_0/\gamma \\
    x' &= \gamma(x-vt) \\
    t' &= \gamma(t-vx/c^2) \\
    u &= \frac{u'+v}{1+u'v/c^2} \\
    f &= f_0 \sqrt{\frac{1-\beta}{1+\beta}} \\
    f &= f_0 \sqrt{1-\beta^2} \\
    \vec{p} &= \gamma m \vec{v} \\
    E &= mc^2 + K = \gamma mc^2 \\
    E^2 &= (pc)^2 + (mc^2)^r
\end{align}

@@

\paragraph{Conductions} (\textbf{pp.1160 ch41 of \cite{book}})
\begin{align}
    P(E) &= \frac{1}{e^{(E-E_F)/kT} +1} 
\end{align}

\section{Anchor}
\label{sec:anchor}
\begin{thebibliography}{1}
    \bibitem{book} Fundamentals of Physics, Extended Edition. Halliday
    \& Resnick.
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

% Resolving pdf version warning
\pdfoptionpdfminorversion=7

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physics-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physics-taper}

@@

This acceleration is due to a net centripetal force on the particle,
with magnitude given by
\begin{align}
    F = \frac{mv^2}{R}
\end{align}
where $m$ is the particle’s mass. The vector quantities $\vec a$ and
$\vec F$ are directed toward the center of curvature of the particle’s
path

\paragraph{First, Second Speed}

\textbf{Circular}: $v = \sqrt{gR}$.

\textbf{Escape Speed} $\frac{1}{2}mv^2 = \frac{GMm}{R}$, so
$v=\sqrt{2GM/R}$.
\begin{defi}[Normal force]
\nomenclature{Normal force}{\nomrefpage.}
A \textbf{normal force} is the force on a body from a surface against
which the body presses.The normal force is always perpendicular to the
surface.
\end{defi}

@@

The linear acceleration $\vec{a}$ of the point \hl{has both tangential and
radial components}.The tangential component is
\begin{align}
    a_t = \alpha r
\end{align}
where $\alpha$ is the magnitude of the angular acceleration (in radians
per second-squared) of the body.The radial component of $\vec{a}$ is
\begin{align}
    a_r = \frac{v^2}{r} = \omega^2 r
\end{align}
If the point moves in uniform circular motion, the period $T$ of
the motion for the point and the body is
\begin{align}
    T = \frac{2\pi r}{v} = \frac{2\pi}{\omega}
\end{align}
\paragraph{Rotational Kinetic Energy and Rotational Inertia} The
kinetic energy $K$ of a rigid body rotating about a fixed axis is
given by
\begin{align}
    K=\frac{1}{2} I\omega^2
\end{align}
in which $I$ is the rotational inertia of the body, defined as
\begin{align}
    I \equiv \int r^2 \diff m
\end{align}
\paragraph{The Parallel-Axis Theorem} 265
\begin{align}
    I = I_{\text{com}} + M h^2
\end{align}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{some-rotational-inertias}.png}
    \caption{Some Rotational Inertias (p. 255 of\cite{book}) }
    \label{fig:some-rotational-inertias}
\end{figure}
\paragraph{Work and Rotational Kinetic Energy} 265
\begin{align}
    W = \int_{\theta_i}^{\theta_f} \tau \diff \theta \\
    P = \frac{\diff W}{\diff t} = \tau \omega
\end{align}

@@

\paragraph{Conduction} (\textbf{pp.498 ch18 of \cite{book}})
\textbf{Thermal expansion}
\begin{align}
    & \Delta L = L \alpha \Delta T \\
    & \Delta V = V \beta \Delta T \\
    & \beta = 3 \alpha \\
    & Q = C \Delta T = cm \Delta T
\end{align}
\begin{defi}[Convection]
\nomenclature{Convection}{\nomrefpage.}
    Convection occurs when temperature differences cause an energy
    transfer by motion within a fluid
\end{defi}
\textbf{Radiation}
\begin{align}
    & P_\text{rad} = \sigma \varepsilon AT^4 \\
    & P_\text{abs} = \sigma \varepsilon AT_\text{evn}^4
\end{align}
\textbf{Conduction}
\begin{align}
    & P_\text{cond} \equiv \frac{Q}{t} = kA \frac{T_H-T_C}{L}  \\
    & P_\text{cond} = A \frac{T_H-T_C}{\sum(L_i/K_i)}
\end{align}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{ch18.heat_cond}.png}
    \caption{pp.494 of \cite{book}}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{ch18.heat_cond_series}.png}
    \caption{pp.495 of \cite{book}}
\end{figure}

\paragraph{Thermal Process} (\textbf{pp.529 ch19 of \cite{book}})
Thermal process:
\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
        No. & Constant Quaility            & Process Type & Special Results  \\
        \hline
        1   & $p$                          & Isobaric     & $Q = nC_p\Delta T;W=p\Delta V$ \\
        2   & $T$                          & Isothermal   & $Q=W=nRT\ln(V_f/V_i); \Delta E_\text{int}=0$  \\
        3   & $pV^\gamma$, $TV^{\gamma-1}$ & Adiabatic    & $W=-\Delta E_\text{int}$ \\
        4   & $V$                          & Isochoric    & $Q=\Delta E_\text{int}=nC_V \Delta T;W=0$ \\
        \hline
    \end{tabular}
\end{table}
In particular, for Adiabatic process (let $K\equiv pV^\gamma$):
\begin{equation}
    W = \int_i^f \frac{K}{V^\gamma}\dd V =
    \frac{p_fV_f-p_iV_i}{1-\gamma}
\end{equation}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{ch18.therm_pro}.png}
\end{figure}
Isothermal Process:
\begin{align}
    W = Nk T \ln\frac{V_f}{V_i}
\end{align}
\textbf{Temperature and Kinetic Energy}
\begin{align}
    & K_\text{avg} = \frac{3}{2}kT \text{, per molecule} \\
    & \lambda = \frac{1}{\sqrt{2} \pi d^2 N/V} 
\end{align}
\paragraph{Molar Specific Heats}
\begin{align}
    & C_V = \frac{3}{2} R \\
    & C_P = C_V + R \\
    & \Delta E_\text{int} = N C_V \Delta T \\
    & C_V = \frac{f}{2} k \\
    & \gamma \equiv \frac{C_P}{C_V}
\end{align}

Even more accurate formulae found in pp. 158 Kardar's book \cite{kardar}:
\begin{align}
    & C_V = \frac{6n-3-r}{2}k_B \\
    & C_p = C_V + k_B
\end{align}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{kardar-C_v-formulae}.PNG}
    \caption{Examples of $n,r,$ and $\gamma$.}
\end{figure}
But these are \hl{\textbf{classical results}}, only valid at high
temperature! (See the refered page for details).

@@

\begin{align}
    & \mathcal{E} = \frac{\diff W}{\diff q}
\end{align}
Cutting Magnetic field lines:
\begin{equation}
    \mathcal{E} = BLv 
\end{equation}
Cutting Magnetic field lines in circular motion:
\begin{equation}
    \mathcal{E} = B(\frac{1}{2}\omega R) R = \frac{1}{2}B\omega R^2
\end{equation}

@@

\paragraph{Radiation} $ $

Electric dipole:
\begin{equation}
    \braket{P} = \frac{\mu_0 p_0^2 \omega^4}{12\pi c}
\end{equation}
where $p_0 = q_0 d$.
Electric quadrupole: $\braket{P} \propto \omega^6$.

(The cases for Magnetic dipole is similarly $\propto \omega^4$)

\textbf{Lienard's generalization of the Larmor formula}
\begin{align}
    P &= \frac{\mu_0 q^2 \gamma^6}{6\pi c}
    \left( a^2- \abs{\frac{\vb{v}\cross\vb{a}}{c}}^2\right) \\
    &\approx \frac{\mu_0q^2}{6\pi c} a^2
\end{align}
(\textbf{pp.448 of \cite{Griffiths_electrod}}).

\begin{center}\noindent\rule{8cm}{0.4pt}\end{center}

@@

\hl{Light waves change phase by $\pi/2$ when they reflect from the surface of
a medium with higher refractive index than that of the medium in which
they are travelling.}

@@

\section{Relativity}
\label{sec:Relativity}

\paragraph{Relativity} (\textbf{pp.1048 ch37 of \cite{book}})
\begin{align}
    \gamma &= \frac{1}{\sqrt{1-\beta^2}} \\
    \Delta t &= \gamma \Delta t_0 \\
    L &= L_0/\gamma \\
    x' &= \gamma(x-vt) \\
    t' &= \gamma(t-vx/c^2) \\
    u &= \frac{u'+v}{1+u'v/c^2} \\
    f &= f_0 \sqrt{\frac{1-\beta}{1+\beta}} \\
    f &= f_0 \sqrt{1-\beta^2} \\
    \vec{p} &= \gamma m \vec{v} \\
    E &= mc^2 + K = \gamma mc^2 \\
    E^2 &= (pc)^2 + (mc^2)^2
\end{align}

\paragraph{Electrodynamics with Relativity} (\textbf{pp.526-531 of
\cite{Griffiths_electrod}})
\begin{equation}
    E = \gamma_0 E_0
\end{equation}
    (Only in the direction perpendicular to
    the motion, without $B$ field)
\begin{align}
    & \bar{E}_x = E_x \\
    & \bar{E}_y = \gamma(E_y - v B_z) \\
    & \bar{E}_z = \gamma(E_z + v B_y) \\
    & \bar{B}_x = B_x \\
    & \bar{B}_y = \gamma(B_y + \frac{v}{c^2} E_z) \\
    & \bar{B}_z = \gamma(B_z - \frac{v}{c^2} E_y)
\end{align}

\section{Quantum Mechanics}
\label{sec:Quantum-Mechanics}

@@

\paragraph{Infinite Square Well} (\textbf{pp.32 of \cite{Griffiths_QM}})

\begin{align}
    & k_n = \frac{n\pi}{a} \\
    & E_n = \frac{\hbar^2 k_n^2}{2m} \\
    & \psi_n = \sqrt{\frac{2}{a}} \sin(k_n x)
\end{align}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/qm_infinite_square_well.PNG}
    \caption{Wave functions}
\end{figure}
\paragraph{Conductions} (\textbf{pp.1160 ch41 of \cite{book}})
\begin{align}
    P(E) &= \frac{1}{e^{(E-E_F)/kT} +1} 
\end{align}

\section{Atomic Physics}
\label{sec:Atomic-Physics}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{electron-state}.PNG}
    \caption{Electron State (from pp.162 of \cite{book_atomic})}
\end{figure}

\paragraph{Standard Model}
Read page 1244 of \cite{book}.
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{Standard-Model-of-Elementary-Particles}.pdf}
    \caption{Standard Model of Elementary Particles (from Wikipedia)}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{Elementary_particle_interactions_in_the_Standard_Model}.png}
    \caption{Elementary particle interactions in the Standard Model
    (from Wikipedia)}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/{Comprehensive-standard-model-particle-chart}.jpg}
    \caption{Comprehensive standard model particle Chart }
\end{figure}
(from \href{http://electron6.phys.utk.edu/phys250/modules/module%206/images/particle_chart.jpg}{Link})

\begin{figure}[H]
    \centering
    \includegraphics[width=1.0\linewidth]{pics/{Elementary-Particles-concept-map}.pdf}
    \caption{Elementary Particles concept Map}
\end{figure}
\section{Positronium}
\label{sec:Positronium}

\href{https://en.wikipedia.org/wiki/Positronium}{Wiki Positronium}.

Energy:
\begin{equation}
    E_n = -\frac{6.8 \text{eV}}{n^2}
\end{equation}
\section{Common Mathematical Formulae}
\label{sec:Common-Mathematical-Formulae}
\begin{align}
    & \frac{1}{1-x} = 1 + x + x^2 + x^3 + \cdots
\end{align}

\section{Anchor}
\label{sec:anchor}
\begin{thebibliography}{1}
    \bibitem{book} Fundamentals of Physics, Extended Edition. Halliday
    \& Resnick.
    \bibitem{book_atomic} Atomic Physics. Fujia, Yang. (Chinese)
    \bibitem{Griffiths_electrod} Introduction to Electrodynamics, 3rd,
    Griffiths.
    \bibitem{Griffiths_QM} Introduction to Quantum Mechanics, 2nd,
    Griffiths.
    \bibitem{kardar} Kardar. Statistical Physics of Particles. (2007)
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}
